---
title: "Data Wrangling with `dplyr`"
subtitle: "FOR 128: Lab 7"
date: "10/17/2024"
format:
  html: default
editor: 
  mode: source
---

## Welcome

Welcome to Lab 7! Today, we'll focus on writing `dplyr` code. In particular, we will both use the verbs individually and "write a sentence" with the verbs by stringing them together with pipes. 

### Learning objectives

-   Use `dplyr` verbs together with pipes. 

### Deliverables (i.e., what to put in the lab drop box)

Upload your rendered PDF (`lab_07.pdf`) **and** Quarto (`lab_07.qmd`) document to the lab drop box. Make sure the Quarto document properly renders to PDF.


## Exercise 0

Load any packages you'll need for this lab below.

```{r}
#| echo: false
#| message: false
library(tidyverse)
```


## Exercise 1

```{r}
#| echo: false
#| message: false
plots <- tibble(
  plot = c(1, 1, 2, 2, 2),
  tree = c(1, 2, 1, 2, 3),
  dbh = c(20.2, 10.4, 5, 18, 10.5),
  logs = c(2, 1, 0.5, NA, 1.5),
  type = c("D", "D", "D", "C", "C"),
  live = c(TRUE, TRUE, TRUE, FALSE, TRUE)
)
```

Create the following dataset and call it `plots`. The resulting tibble should look like this when printed:

```{r}
plots
```


## Exercise 2

Write some code to figure out the follow features of `plots`:

  a. How many rows and columns?
  b. What are the column names?
  c. What is the data type of each column?
  d. Are there any `NA` values? If so, in which column?
  
## Exercise 3

Use a `dplyr` function to print all trees in plot 2. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2)
```

### Exercise 4

Use a `dplyr` function to print all trees in plot 2 that have dbh less than or equal 10. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2 & dbh <= 10)
```

### Exercise 5

Use a `dplyr` function to print the tree with the largest dbh. 

```{r}
#| echo: false
#| message: false
plots %>% slice_max(dbh)
```

### Exercise 6

Use a series of piped `dplyr` functions to find the largest dbh tree on plot 2.

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2) %>% slice_max(dbh)
```

### Exercise 7

Use a series of piped `dplyr` functions to find the largest dbh live tree on plot 2. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2 & live) %>% slice_max(dbh)
```

### Exercise 8

Use a series of piped `dplyr` functions to find the largest dbh dead tree on plot 2. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2 & !live) %>% slice_max(dbh)
```

### Exercise 9

Use a series of piped `dplyr` functions to find the largest dbh live tree on plot 2 of type D. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 2 & live & type == "D") %>% slice_max(dbh)
```

### Exercise 10

Use a series of piped `dplyr` functions to find the smallest dbh tree on plot 1. 

```{r}
#| echo: false
#| message: false
plots %>% filter(plot == 1) %>% slice_min(dbh)
```

### Exercise 11

Use a `dplyr` function to add a new column to `plots` to hold each tree's basal area (ft$^2$). This new column should be called `ba` with values equal to `0.005454*dbh^2` (assuming dbh is in inches).

```{r}
#| echo: false
#| message: false
plots <- plots %>% mutate(ba = 0.005454*dbh^2)
plots
```

### Exercise 12

Use a `dplyr` function to move your newly created column `ba` to between the `dbh` and `logs` columns.

```{r}
#| echo: false
#| message: false
plots <- plots %>% relocate(ba, .after="dbh")
plots
```

### Exercise 13

Use a series of piped `dplyr` functions to compute the mean dbh for trees on plots 1 and 2. Note, I called my mean `mean_dbh`.

```{r}
#| echo: false
#| message: false
plots %>% group_by(plot) %>% summarise(mean_dbh = mean(dbh))
```

### Exercise 14

Use a series of piped `dplyr` functions to compute plot specific mean dbh and logs for trees. Exclude `NA` values from the mean calculations (hint, use the `na.rm` argument in `mean()`). Note, I called my mean `mean_dbh` and `mean_logs`.

```{r}
#| echo: false
#| message: false
plots %>% group_by(plot) %>% summarise(mean_dbh = mean(dbh), mean_logs = mean(logs, na.rm=TRUE))
```

### Exercise 15

Use a series of piped `dplyr` functions to compute plot specific mean dbh and logs for *live* trees. Note, I called my mean `mean_dbh` and `mean_logs`. Why did only plot 2 mean dbh change from the your solution to Exercise 14?

```{r}
#| echo: false
#| message: false
plots %>% filter(live) %>% group_by(plot) %>% summarise(mean_dbh = mean(dbh), mean_logs = mean(logs, na.rm = TRUE))
```

### Exercise 16

Sort `plots` by increasing plot number and increasing dbh within plot. 

```{r}
#| echo: false
#| message: false
plots %>% arrange(plot, dbh)
```

### Exercise 17

The type column holds values "D" and "C" which stand for deciduous and conifer, respectively. Use `mutate()` and the `case_when()` function to change values in the type column from "D" and "C" to "deciduous" and "conifer". 
```{r}
#| echo: false
#| message: false
plots <- plots %>% mutate(type = case_when(
  type == "D" ~ "deciduous",
  type == "C" ~ "conifer"
))
```
```{r}
plots
```

### Exercise 18

Use a series of piped `dplyr` functions to compute type specific mean dbh and logs. More specifically, I want you to use a grouped `summarize()`, where you group by type. Note, I called my mean `mean_dbh` and `mean_logs`.

```{r}
#| echo: false
#| message: false
plots %>% group_by(type) %>% summarise(mean_dbh = mean(dbh, na.rm = TRUE), mean_logs = mean(logs, na.rm = TRUE))
```

### Exercise 19

Use a series of piped `dplyr` functions to count the number of trees by type. Hint, use the `n()` within a grouped `summarize()`. I called my count `n_trees`.

```{r}
#| echo: false
#| message: false
plots %>% group_by(type) %>% summarise(n_trees = n())
```

### Exercise 20

Use a series of piped `dplyr` functions to print the trees with the largest basal area within each plot. 

```{r}
#| echo: false
#| message: false
plots %>% group_by(plot) %>% slice_max(ba)
```


# Wrap up

Congratulations! You've made it to the end of Lab 7. Make sure to render your final document and submit both the .pdf and .qmd file to D2L.
