---
title: "Writing Functions"
subtitle: "FOR 128: Lab 5"
author: "Insert Your Name Here"
date: "10/2/2024"
format:
  pdf: default
editor: 
  mode: source
---

## Welcome

Welcome to Lab 5! Today, we'll focus on writing our own custom functiuons. 

### Learning objectives

-   Get comfortable writing functions in `R`.
-   Specify multiple functionarguments and function argument defaults.
-   Create informative error messages with input checks. 
-   Use conditional statements within functions. 

### Deliverables (i.e., what to put in the lab drop box)

Upload your rendered PDF (`lab_05.pdf`) **and** Quarto (`lab_05.qmd`) document to the lab drop box. Make sure the Quarto document properly renders to PDF.

### Collaborator(s)

List any collaborators you worked with below. 



## Exercise 1

Write a function that takes the absolute value of the difference of the mean and median of a vector. You function should allow the user to choose to remove NAs or not. Make sure to add a check in your function that the vector is numeric, as well. 

```{r}

```


## Exercise 2

Do Exercise 5.1 from IFDAR

```{r}

```


## Exercise 3

Do Exercise 5.2 from IFDAR

```{r}

```


## Exercise 4

Do Exercise 5.3 from IFDAR

```{r}

```


## Exercise 5

Do Exercise 5.4 from IFDAR

```{r}

```


## Exercise 6

Do Exercise 5.5 from IFDAR

```{r}

```


## Exercise 7

Do Exercise 5.6 from IFDAR

```{r}

```


# Wrap up

Congratulations! You've made it to the end of Lab 5. Make sure to render your final document and submit both the .pdf and .qmd file to D2L.
