---
title: "FOR 128: Lab 1"
author: "Insert Your Name Here"
date: "8/29/2024"
format:
  pdf: default
---

## Welcome

Welcome to Lab 1! Today, we'll get started with a few data science tools (`R`, RStudio, and Quarto), and learn to use `R` as a calculator.

### Learning objectives

-   Make sure `R` and RStudio are set up correctly.
-   Set up folders on your computer for saving your work.
-   Begin working with Quarto (.qmd) documents.
-   Learn to save and render your work with Quarto.
-   Start using `R` as a calculator.

### Deliverables (i.e., what to put in the lab drop box)

Upload your rendered PDF (`lab_01.pdf`) **and** Quarto (`lab_01.qmd`) document to the lab drop box. Make sure the Quarto document properly renders to PDF.

# Part 1: Setting up your workspace

## Exercise 1

Create a folder on your computer where you'll store all your work for this course. We do not suggest putting this on your Desktop. A better place would be your Documents folder or your home directory. Make sure to give the folder an informative name like `for128` or `for128_f24` or something else. Put this file (`lab_01.qmd`) in your new folder.

Now, run the following code chunk (press the green arrow on the far right of the code chunk). The output should include this file. What do you think the `list.files()` function does in this case? Please answer between the rows of asterisks (stars).

```{r}
list.files()
```

***





***

## Exercise 2

Create a subfolder called `data` in the folder you just created in Exercise 1. For now, we'll leave this directory empty, but we will use it for future labs.

Now, run the following code chunk. The output should include the lab 2 data file (`lab_02.txt`). What do you think specifying the `path` argument in the `list.files()` function did? Please answer between the rows of asterisks (stars).

```{r}
list.files(path = "data/")
```

***





***

## Exercise 3

Render this document by clicking "Render" towards the top of the RStudio window. What does it look like? Does the output of your code from Exercises 1 and 2 look the same? If it doesn't, why do you think that is the case?

Note: You can answer this Exercise just with words, so there is no need for a code chunk for this Exercise.

***





***



# Part 2: `R` as a calculator

## Exercise 4

Do Exercise 2.1 from IFDAR in the following code chunk.

***

```{r}
# put your code on the following line

```

***

## Exercise 5

Do Exercise 2.2 from IFDAR in the following code chunk.

***

```{r}

```

***

## Exercise 6

Do Exercise 2.3 from IFDAR. Note, you'll have to make your own code chunk. 

Hint: on Windows and Linux you can create a new code chunk with control + alt + I and on macOS you can make a new code chunk with command + alt + I. Otherwise, you can create a new code chunk by clicking the code dropdown menu in RStudio and then clicking Insert Chunk. 

***





***


## Exercise 7

Do Exercise 2.4 from IFDAR. Note, you'll have to make your own code chunk.

***





***


# Wrap up

Congratulations! You've made it to the end of Lab 1. Make sure to render your final document and submit both the .pdf and .qmd file to D2L.
